---
title: 'Statistical report: Use of the inhibitors of sodium-glucose cotransporter 2 in patients with heart failure'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    dev: cairo_pdf
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
   - \usepackage{subfig}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB4, https://kiheartfailure.github.io/shfdb4/, v 4.0.1. 

## Inclusion/exclusion criteria

```{r, child = "./src/flowchart.Rmd"}

```

*if a patient dies immediately after the index visit he/she will not collect a drug prescribed at the index visit
and thereby be defined as a digoxin user. Throughout, the index date is moved 14 days forward.

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`.  

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

## Created variables 

Prevalent users of SGLT2i are defined as patients with a dispensed prescription for SGLT2i
prior to 5 months. 

```{r atc}
default_kable(metalm, caption = "Treatments from DDR")
```

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1}. The primary outcome, first heart failure hospitalization/CVD,
was included as the Nelson-Aalen estimator. SGLT2i use is not included in the imputation model.  

## Propensity scores and matching

A propensity score for treatment with SGLT2i was estimated for each patient with 
logistic regression for each of the 10 imputed datasets using the variables 
indicated in Table \ref{tab:tab1}. 1:1 matching without replacement [@match] 
was thereafter performed on the average of the resulting 10 ps [@psmatch]. Matching was 
allowed if the propensity score differed by 0.01 or less. The ability of the ps 
matching to balance the baseline characteristics was assessed by 
standardized mean differences.  

1:1 has deemed the best option when taking into account the number of patients retained and balance between groups.
Other matching ratios: `r matchingn`.

\clearpage

## SGLT2i over time

The denominator consists of persons that at the 
middle of the month were present in SHFDB4 with the specified inclusion/exclusion criteria 
(so included in SwedeHF prior to 20xx-xx-15 but not dead/emigrated as of 20xx-xx-15). 
The numerator consists of the number of patients of the above with at least one SGLT2i 
dispension during that year.  

```{r, child = "./src/overtime.Rmd"}

```

\clearpage

\blandscape

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}

```

\elandscape

## Associations with SGLT2i use

Odds ratios are estimated using logistic regression with SGLT2i use as outcome. 
Outliers were investigated with Cook's distance and multicollinearity 
with the variance inflation factor.
=> XX. 

```{r, child = "./src/ortab.Rmd"}

```


\clearpage

## Association SGLT2i vs outcomes

The following outcomes are considered: 

- Cardiovascular mortality/first HF hospitalization (primary endpoint)
- Cardiovascular mortality
- First HF hospitalization
- Total HF hospitalization
- All-cause mortality
- First kidney hospitalization

Incidence per 1000 py was calculated with 95% Poisson confidence intervals. 

Time to first events were presented with cumulative incidence curves and 
the mean cumulative function (MCF) was used to depict the recurrent events. 

Cox proportional hazards regressions were used to evaluate the association 
between SGLT2i and time to first event, partly crude and partly adjusted for variables indicated in 
Table \ref{tab:tab1}. The variables were selected based on clinical relevance. 
The association was also evaluated in the ps matched cohort 
(so also adjusted for variables in Table \ref{tab:tab1}). 
In the matched cohort the matched pairs were modelled using a frailty term with a gaussian distribution. 

As a consistency analysis the time to first events were modelled using a sub-distributional hazards model [@fg] 
where death was treated as a competing event. 

Recurrent events were model using a negative binomial regression 
including the log of time as an offset in the model adjusting as above. For the matched cohort a 
generalized linear mixed-effects model (GLMM) for the negative binomial family 
with matched pair as a random effect was used. 

Data were censored at 2021-12-31, death from other causes than the event or emigration from Sweden. 

Overall cohort: The median (min-max) follow-up is 
`r rsdata %>% summarise(med = fn(median(sos_outtime_death / 365.25 * 12), dig = 1),
                                             min = fn(min(sos_outtime_death / 365.25 * 12), dig = 1),
                                             max = fn(max(sos_outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r rsdata %>% summarise(sumpy = fn(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


Matched cohort: The median (min-max) follow-up is 
`r matchrsdata %>% summarise(med = fn(median(sos_outtime_death / 365.25 * 12), dig = 1),
                                             min = fn(min(sos_outtime_death / 365.25 * 12), dig = 1),
                                             max = fn(max(sos_outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r matchrsdata %>% summarise(sumpy = fn(sum(sos_outtime_death) / 
                                   365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary outcome. 
Possible outliers were visually inspected by plotting the dfbetas. 
=> XXXX. 

\clearpage

### Crossover

In the main analysis SGLT2i use is considered according to the ITT principle 
(SGLT2i at baseline). As a consistency analysis SGLT2i was included as a 
time-dependent variable in the model for the time to first event outcomes. 

In patients treated with SGLT2i at baseline the patient was assumed to be on SGLT2i 
until LAST dispension + 3 months or end of follow-up - 5 months independent of how 
long time between dispensions. In untreated patients, crossover was defined 
as a new SGLT2i dispension in the The National Prescribed Drug Register and 
thereafter in the same manner as for patients treated with SGLT2i at baseline. 
That means that patients with SGLT2i at baseline were allowed to cross 
over max once (Yes-No) and patients without SGLT2i at baseline were 
allowed to cross over max twice (No-Yes-No). 

No of cross-overs (until death): 

`r default_kable(inner_join(matchrsdata,
  matchcross_death %>%
    group_by(lopnr) %>%
  slice(2) %>%
    ungroup() %>%
    select(lopnr),
  by = "lopnr"
) %>%
  count(sos_ddr_sglt2), 
  scale_down = FALSE)`

```{r, child = "./src/outtab.Rmd"}

```

\clearpage

```{r, child = "./src/km.Rmd"}

```

\clearpage

### Subgroups

Performed using interaction effects on the patients in the matched cohort. 
Note that there is missing data for eGFR and diabetes (type I excluded) so the matching (and thereby the adjustment) 
may therefore not be complete. 

```{r, child = "./src/outtabsub.Rmd"}

```

\clearpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/sglt2ihf. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
