---
title: 'Statistical report: Use of the inhibitors of sodium-glucose cotransporter 2 in patients with heart failure'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2: 
    dev: cairo_pdf
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{subfig}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
---

\clearpage

\listoftables
\listoffigures

\clearpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SwedeHF combined with national registers, similar to SHFDB4, 
https://kiheartfailure.github.io/shfdb4/, v 4.0.1, but with longer follow-up. 

## Inclusion/exclusion criteria

```{r, child = "./src/flowchart.Rmd"}

```

*if a patient dies immediately after the index visit he/she will not collect a drug prescribed at the index visit
and thereby be defined as a SGLT2i user. 

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`.  

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

For the supplementary analysis excluding patients with Type I diabetes, eGFR < 20 and dialysis: 

First patient in: `r min(rsdata$shf_indexdtm[rsdata$subset])` and last patient in: 
`r max(rsdata$shf_indexdtm[rsdata$subset])`.  

The median age (IQR) is `r rsdata %>% filter(subset) %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% filter(subset) %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    
  
## Created variables 

```{r atc}
default_kable(metalm, caption = "Treatments from Prescribed Drug Register")
```

```{r com}
default_kable(outcommeta, caption = "Comorbidities from NPR")
```

```{r cci}
default_kable(ccimeta, caption = "Charlson Comorbidity Index from NPR")
```

Socio-economic variables from LISA are on a yearly basis and set to the year prior 
to registration in SwedeHF. When data was extracted data for 2021 was not available 
and therefore patients with index year 2022 receive values from 2000. 

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1}. SGLT2i use is not included in the imputation model.  

\clearpage

## SGLT2i over time

In Figure \ref{fig:overtime} the denominator consists of persons that at each half year time period
were present in SwedeHF with the specified inclusion/exclusion criteria 
(so included in SwedeHF prior to 202x-x1/7-01 but not dead/emigrated as of 202x-1/7-01). 
The numerator consists of the number of patients of the above with at least one SGLT2i 
dispension during that halfyear time period. 

In Figure \ref{fig:overtime2} the proportion of patients that, within 5 months 
prior up until 14 days after their registration in SwedeHF, collect an SGLT2 prescription is presented. 

```{r, child = "./src/overtime.Rmd"}

```

\clearpage

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}

```

\clearpage

```{r, child = "./src/tab1_rowpercent.Rmd"}

```

\clearpage 

## Variation in SGLT2i use across regions

The region is "sjukvårdsregion" that the hospital/primary care centre belongs to. 

```{r, child = "./src/regional.Rmd"}

```

\clearpage 

## Variation in SGLT2i use across centres

```{r, child = "./src/mor.Rmd"}

```

In order to evaluate the difference in likelihood of being prescribed SGLT2i at two randomly 
selected centres the median odds ratio (MOR)[@mor] was calculated for a model 
including the intercept and centre as a random variable. 

The MOR was `r fn(MOR$vcov, 2)`. 

The number of centres with $\le$ 10 patients was `r kollcenter %>% filter(n10 == "<=10") %>% pull(np)`.

\clearpage 

## Associations with SGLT2i use

Odds ratios are estimated using logistic regression with SGLT2i use as outcome. 
Outliers were investigated with Cook's distance and multicollinearity 
with the variance inflation factor and no action was deemed necessary.

```{r, child = "./src/ortab.Rmd"}

```

\clearpage

## Associations with SGLT2i use in subgroups

Odds ratios are estimated using logistic regression with SGLT2i use as outcome. 
The subgroup variable and the other respective variable were included in the model 
using an interaction thereby rendering the interpretation of a subgroup analyses 
but with the advantage of obtaining a p-value. Note therefore that also the adjusted 
estimates are obtained from different models. 

```{r, child = "./src/ortab_int.Rmd"}

```

\clearpage

## Discontinuation in SGLT2i

Discontinuation is considered to occur if there is > 5 months gap between pill dispensations. If this is the case, the discontinuation date is set to 3 months after the last dispensation. The % discontinued at 6 months, 1 year and 1.5 years after index is estimated with the Kaplan-Meier method were patients are censored at last follow-up 
(`r global_endfollowup`), death or emigration.  

```{r, child = "./src/disc.Rmd"}

```

\clearpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/sglt2ihf. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
